---
title: 'Exploring the relationship between altmetrics and traditional measures of dissemination in health professions education'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library a few packages in the for the exploratory analysis and models.
```{r, warning = FALSE, message = FALSE}
library(ggplot2)
library(gridExtra)
library(psych)
library(MASS)
```

Point your working directory as appropriate.
```{r}
setwd("E:/USU/LEROUX/RESEARCH/TWEETS")
```

Read in the CSV file.
```{r}
dta <- read.csv(file = "altmetrics_data.csv", header = TRUE)
```

You should have 3,243 obserations and 17 variables (let's just check)
```{r}
dim(dta)
```

Examining citation-related data.
```{r}
plot(density(dta$cited))

########## MODEL # 1 - NEGATIVE BINOMIAL ##########
## no interaction
## of note: the variable labeled 'open' refers to publicly accessible
fit1 <- glm.nb(formula = cited ~ open + references + bloggers + tweets + news_outlets + 
                                 facebook + mendeley + impact + time, data = dta)
summary(fit1)

## model with interaction (used in the manuscript)
fit2 <- glm.nb(formula = cited ~ open + references + bloggers + tweets + news_outlets + 
                                 facebook + mendeley + impact + time + open:impact, data = dta)
summary(fit2)
cbind(rate = exp(fit2$coefficients), CI = exp(confint(fit2)))
```

Now for the access counts.
```{r, message = FALSE}
plot(density((dta$updated_usage)))

########## MODEL # 2 - OLS ##########
## no interaction
fit3 <- lm(formula = updated_usage ~ open + references + bloggers + tweets + news_outlets + 
                 facebook + mendeley + impact + time , data = dta)
summary(fit3)

## model with interaction (used in the manuscript)
fit4 <- lm(formula = updated_usage ~ open + references + bloggers + tweets + news_outlets + 
                 facebook + mendeley + impact + time + open:impact, data = dta)
summary(fit4)
cbind(estimate = fit4$coefficients, ci = confint(fit4))
```

Some plots of interest (maybe).
```{r}
smoothScatter(x = dta$updated_usage[dta$open == 0], y = dta$impact[dta$open == 0])
points(x = dta$updated_usage[dta$open == 1], y = dta$impact[dta$open == 1], col = "red")
```

We cannot really see a whole lot with the above plot. Let's turn to ggplot2 to get a bit more detail.
```{r, warning = FALSE}
commonTheme = list(labs(color = "Density", fill = "Density",
                        x = "Usage", y = "Impact Factor",
                        theme_bw(),
                        theme(legend.position = c(0,1),
                              legend.justification = c(0,1))))

ggplot(data = dta, aes(x = dta$updated_usage,y = dta$impact)) + 
  geom_density2d(aes(color = ..level..)) +
  scale_color_gradient(low = "green", high = "red") + 
  geom_point() + commonTheme
```

Okay, the above plot was a little better -- let's stratify by publicly accessible for more detail (open versus closed).
```{r, warning = FALSE}
dta_closed <- dta[dta$open == 0 , ]
dta_open <- dta[dta$open == 1 , ]

plot1 <- ggplot(data=dta_closed, aes(x = updated_usage,y = impact)) + ggtitle("Closed Articles") + expand_limits(y = c(0.5, 4.2)) +
  stat_density2d(aes(fill = ..level..,alpha=..level..),geom = 'polygon', color = 'black') + 
  scale_fill_continuous(low = "green",high = "red") +
  geom_smooth(method = lm,linetype=2,colour = "red",se = F) + 
  guides(alpha = "none") +
  geom_point() + commonTheme

plot2 <- ggplot(data=dta_open, aes(x = updated_usage,y = impact)) + ggtitle("Open Articles") + expand_limits(y = c(0.5 ,4.2)) +
  stat_density2d(aes(fill = ..level..,alpha = ..level..), geom = 'polygon', color = 'black') + 
  scale_fill_continuous(low = "green", high = "red") +
  geom_smooth(method = lm,linetype = 2, colour = "red", se = F) + 
  guides(alpha = "none") +
  geom_point() + commonTheme

grid.arrange(plot1, plot2, ncol = 2)
```